# Building a REDCap database {#redcap-data}

I will try to provide examples for two cases: 

## Creating a database {#redcap-scratch}

Lessons learned...


## Setting up a database from a dataset {#redcap-dataset}

Below will follow steps necessary to get a data base set up based on a data set provided.

### Create a data dictionary

Copy and run the following to create a DataDictionary draft.

Change the `var_name` vector to the variable names of your own project and the same for `instrument`. To fill out the rest, open the .ods file in a visual editor as LibreOffice go manual.

```{r redcap-dd-1}
require(readODS)

var_name <- c("record_id",colnames(stRoke::talos))
instrument = c("incl","incl","follow1","follow6","incl","incl","incl")

(df <- data.frame(Variable...Field.Name=var_name,
           Form.Name=instrument,
           Field.Type = "text",
           Choices..Calculations..OR.Slider.Labels = "",
           Field.Label = ""))

write_ods(df,
          path = "dd_draft.ods")
```

Have a look at the demonstration file below for inspiration.

```{r redcap-dd-example}
head(read.csv("./redcap/REDCapDataDictonaryDemo.csv"))
```

_Next steps are coming_
